---
title: "Updated development"
author: "Hannah Gurholt"
date: "2024-11-11"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

##Loading Packages
```{r setup, loading packages}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
library(lubridate)
library(nlme)
##install.packages("sjPLot")
library(sjPlot)
library(survival)
##install.packages("coxme")  # Install the package
library(coxme)  
library(ggeffects)
library(dplyr)
library(ggplot2)
#install.packages("emmeans")
library(emmeans)
#install.packages("lme4")
library(lme4)
#install.packages("nlme")
library(nlme)
library(multcomp)
#install.packages("multcompView")
library(multcompView)
#install.packages("survival")
#install.packages("survminer")
library(survival)
library(survminer)
#install.packages("knitr")
library(knitr)
#install.packages("patchwork")
library(patchwork)
library(car)
library(lavaan)
library(fastDummies)
library(kableExtra)
#install.packages("officer")
library(officer)
library(flextable)
library(parameters)
#install.packages("ggcorrplot")  # if not installed
library(ggcorrplot)
#install.packages("pheatmap")
library(pheatmap)
#install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
library(survival)
library(survminer)
library(GGally)

```

## Data entry and processing
```{r, load data}
raw_mon <- read.csv("/Users/hannahgurholt/Desktop/monALAN2023/data_raw/Data Analysis Trial Raw data 8.31.23.csv")

summary(raw_mon)

```

##Cleaning the data 
```{r, name the data frame and convert variables through the mutate function}
data_mon <- raw_mon %>%
  mutate(Hatch.Date = mdy(Hatch.Date), # Dates
         Day.7.date = mdy(Day.7.date),
         Date.of.pupation = mdy(Date.of.pupation),
         Date.of.eclosion = mdy(Date.of.eclosion),
         Date.of.death = mdy(Date.of.death)) %>%
  mutate(Cohort = factor(Cohort), # Factors
         Species = factor(Species),
         Stage.at.move = factor(Stage.at.move),
         Sex = factor(Sex),
         OE.parasite = factor(OE.parasite),
         CohortCage = factor(paste(Cohort, Cage, sep="."))) %>%
  mutate(treatment = relevel(factor(substr(Cage, 1, 2)), ref = "ML"),
         Cage = factor(Cage)) %>%
  mutate(time.to.pupa = day(days(Date.of.pupation - Hatch.Date)),
         time.to.death =  day(days(Date.of.death - Hatch.Date)),
         development.time = day(days(Date.of.eclosion - Hatch.Date)),
         eclosed = ifelse(is.na(Date.of.eclosion), 'died', 'eclosed'),
         status = ifelse(is.na(Date.of.pupation), 'died', 'pupated')) %>%
  mutate(growth.rate = (Day.7.Mass - Initial.Mass)/6) %>%
  filter(Individual != "" & !is.na(treatment) & treatment != "" & Species != "") %>%
  mutate(treatment = droplevels(treatment, exclude = ""),
         Species = droplevels(Species, exclude = ""),
         Sex = droplevels(Sex, exclude = ""))

levels(data_mon$Sex)
levels(data_mon$treatment)
summary(data_mon)

parse_date_time(raw_mon$Date.of.pupation, "%m%d%y")
unique(raw_mon$Date.of.pupation)


cohort_labels <- c(
  "1" = "July 2023",
  "2" = "August 2023",
  "3" = "September 2023",
  "4" = "August 2024"
)

sex_labels<- c(
  "M" = "Male",
  "F" = "Female"
)

```

## Growth rate as a function of ALAN
```{r, model growth, post hoc test, and figure}
# Start with a simpler model to look at all of the possible interactions that could be going on
model_growth <- lme(growth.rate ~ treatment + Cohort + Species + Sex,
                    random = ~ 1 | CohortCage,
                    data = data_mon,
                    na.action = na.omit)


summary(model_growth)
kable(summary(model_growth)$tTable)

#Run and Anova
#anova(model_growth)
anova(model_growth, type="marginal")
#Anova(model_growth, type= "II")
#Anova(model_growth, type= "III")
#don't understand why the type 3 anova won't work for this one 

# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_growth <- emmeans(model_growth, pairwise ~ treatment | Species + Cohort + Sex)

# Perform the pairwise comparisons with Tukey adjustment
tukey_results_growth <- pairs(emmeans_results_growth, adjust = "tukey")

kable(summary(tukey_results_growth)$Table)

```

###Figure 1A
##Graphing for growth rate
```{r, growth rate figure}
#making predictions
growth_pred <- emmeans(model_growth, specs = ~ treatment + Species) %>%
  as.data.frame()

#summary of predictions
growth_pred_summary <- growth_pred %>%
  group_by(Species, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))


growth_plot<- ggplot(growth_pred_summary, aes(x = treatment, y = 1000 * emmean, color = Species)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Lines connecting means
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  
  # Error bars
  geom_errorbar(aes(ymin = 1000 * lower.CL, ymax = 1000 * upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Custom colors for species
  scale_color_manual(name = "Species",
                     values = c("inc" = "skyblue3", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Growth rate (mg/day)"))) +
  xlab("Treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
  # Y-axis zoom
  coord_cartesian(ylim = c(5, 30)) +
  
    labs(tag = "A") +

  # Plot aesthetics
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),        # Remove all panel borders
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.line.y.right = element_blank(),   # Remove right y-axis line
    axis.line.x.top = element_blank(),     # Remove top x-axis line
    legend.position = "none"   , 
        
    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)# Hide legend
  )


#now inlcuding some raw jittered points 

growth_plot <- ggplot(growth_pred_summary, aes(x = treatment, y = 1000 * emmean, color = Species)) +
  
  # Raw data points (behind)
  geom_jitter(
    data = data_mon,
    aes(x = treatment, y = 1000 * growth.rate, color = Species),
    alpha = 0.2,             # transparency for visibility behind means
    size = 1.8,              # small circles
    shape = 16,              # filled circles
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
  ) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Lines connecting means
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  
  # Error bars
  geom_errorbar(
    aes(ymin = 1000 * lower.CL, ymax = 1000 * upper.CL),
    width = 0.1,
    position = position_dodge(width = 0.5)
  ) +
  
  # Custom colors for species
  scale_color_manual(
    name = "Species",
    values = c("inc" = "deepskyblue4", "syr" = "red3"),
    labels = c(expression(italic("A. incarnata")),
               expression(italic("A. syriaca")))
  ) +
  
  # Axis labels
  ylab(expression(paste("Growth rate (mg/day)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
  # Y-axis zoom
  coord_cartesian(ylim = c(5, 30)) +
  
  labs(tag = "A") +
  
  # Plot aesthetics
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    axis.line.y.right = element_blank(),
    axis.line.x.top = element_blank(),
      legend.position = c(0.3, 0.93),
    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)
  )


growth_plot
```

##larval duration as a function of ALAN
```{r, model larval duration, and post hoc testing and plot}
#no signficiant interactions 

model_larvalduration <- lme(Larval_duration ~ treatment + Sex + Cohort + Species,
                    random =~ 1|CohortCage,
                    data = data_mon,
                    na.action = na.omit)

summary(model_larvalduration)
kable(summary(model_larvalduration)$tTable)

anova(model_larvalduration, type="marginal")
#Run and Anova
#Anova(model_larvalduration, type= "III")

#Anova(model_larvalduration, type= "II")

#Anova(model_larvalduration)
#anova(model_larvalduration)

# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_larvalduration <- emmeans(model_larvalduration, ~ treatment + Species + Cohort + Sex)

# Perform the pairwise comparisons with Tukey adjustment
tukey_results_larvalduration <- pairs(emmeans_results_larvalduration, adjust = "tukey")
print(tukey_results_larvalduration)

```

###Figure 1B
##Larval duration predictions and plot
```{r}

#making predictions
larval_pred <- emmeans(model_larvalduration, specs = ~ treatment + Species) %>%
  as.data.frame()

#summary of predictions
larval_pred_summary <- larval_pred %>%
  group_by(Species, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))


larval_duration_plot <- ggplot(larval_pred_summary, aes(x = treatment, y = emmean, color = Species)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Lines connecting means
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  
  # Error bars
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Custom colors for species
  scale_color_manual(name = "Species",
                     values = c("inc" = "skyblue3", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Larval duration (days)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
    
  # Y-axis zoom
  coord_cartesian(ylim = c(15, 21)) +
  
    labs(tag = "B") +

  # Plot aesthetics
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),       
    axis.line = element_line(color = "black"),  
    axis.line.y.right = element_blank(),   
    axis.line.x.top = element_blank(),     
    legend.position = "none"    , 
    
    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)
  )


#now with the raw jittered data points 
larval_duration_plot <- ggplot(larval_pred_summary, aes(x = treatment, y = emmean, color = Species)) +
    
  # Raw data points (behind)
  geom_jitter(
    data = data_mon,
    aes(x = treatment, y = Larval_duration, color = Species),
    alpha = 0.2,             # transparency for visibility behind means
    size = 1.8,              # small circles
    shape = 16,              # filled circles
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
  ) +
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Lines connecting means
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  
  # Error bars
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Custom colors for species
  scale_color_manual(name = "Species",
                     values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Larval duration (days)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
    
  # Y-axis zoom
  coord_cartesian(ylim = c(15, 21)) +
  
    labs(tag = "B") +

  # Plot aesthetics
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),       
    axis.line = element_line(color = "black"),  
    axis.line.y.right = element_blank(),   
    axis.line.x.top = element_blank(),     
    legend.position = "none"    , 
    
    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)
  )

larval_duration_plot
```

##pupal duration as a function of ALAN
```{r pupal duration model, post hoc, and figure}

#no significant interactions so we will revise the model

model_pupalduration <- lme(Pupal_duration ~ treatment + Sex + Cohort + Species,
                    random =~ 1|CohortCage,
                    data = data_mon,
                    na.action = na.omit)

summary(model_pupalduration)
kable(summary(model_pupalduration)$tTable)

#Run an Anova
anova(model_pupalduration, type="marginal")
Anova(model_pupalduration, type= "III")



# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_pupalduration <- emmeans(model_pupalduration, ~ treatment + Species + Cohort + Sex)

# Perform the pairwise comparisons with Tukey adjustment
tukey_results_pupalduration <- pairs(emmeans_results_pupalduration, adjust = "tukey")
print(tukey_results_pupalduration)

```

###Figure 1C
#Predicting pupal duration and plotting it 
```{r}
# Making predictions for pupal duration
pupal_pred <- emmeans(model_pupalduration, specs = ~ treatment + Species) %>%
  as.data.frame()

# Summary of predictions
pupal_pred_summary <- pupal_pred %>%
  group_by(Species, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))


pupal_duration_plot <- ggplot(pupal_pred_summary, aes(x = treatment, y = emmean, color = Species)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Lines connecting means
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  
  # Error bars
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Custom colors for species
  scale_color_manual(name = "Species",
                     values = c("inc" = "skyblue3", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Pupal duration (days)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
  # Y-axis zoom (adjust as needed)
  coord_cartesian(ylim = c(11, 17)) +
  
  # Plot aesthetics
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),       
    axis.line = element_line(color = "black"),  
    axis.line.y.right = element_blank(),   
    axis.line.x.top = element_blank(),     
    legend.position = "none"               
  )


pupal_duration_plot <- ggplot(pupal_pred_summary, aes(x = treatment, y = emmean, color = Species)) +
  
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  scale_color_manual(name = "Species",
                     values = c("inc" = "skyblue3", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  ylab(expression(paste("Pupal duration (days)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  coord_cartesian(ylim = c(11, 17)) +

  labs(tag = "C") +

  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    axis.line.y.right = element_blank(),
    axis.line.x.top = element_blank(),

    # Legend positioning and style
    legend.position = c(0.95, 0.05),      # bottom right inside plot (x=right, y=bottom)
    legend.justification = c("right", "bottom"), 
    legend.background = element_blank(),  # No background (no box)
    legend.key = element_blank(),         # No key boxes around symbols
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),

    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)
  )


#now adding raw jittered data points 
pupal_duration_plot <- ggplot(pupal_pred_summary, aes(x = treatment, y = emmean, color = Species)) +
    # Raw data points (behind)
  geom_jitter(
    data = data_mon,
    aes(x = treatment, y = Pupal_duration, color = Species),
    alpha = 0.2,             # transparency for visibility behind means
    size = 1.8,              # small circles
    shape = 16,              # filled circles
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
  ) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  geom_line(aes(group = Species), position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1, position = position_dodge(width = 0.5)) +
  scale_color_manual(name = "Species",
                     values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  ylab(expression(paste("Pupal duration (days)"))) +
  xlab("Light treatment") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  coord_cartesian(ylim = c(11, 17)) +

  labs(tag = "C") +

  theme_bw(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    axis.line.y.right = element_blank(),
    axis.line.x.top = element_blank(),

    # Legend positioning and style
    legend.position = "none",      # bottom right inside plot (x=right, y=bottom)
    legend.justification = c("right", "bottom"), 
    legend.background = element_blank(),  # No background (no box)
    legend.key = element_blank(),         # No key boxes around symbols
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),

    plot.tag.position = c(0.02, 0.98),
    plot.tag = element_text(size = 18)
  )

pupal_duration_plot
```

###Figure 1
#Putting together growth, larval duration, and pupal duration for Figure 1
```{r}
Figure_1<- growth_plot + larval_duration_plot + pupal_duration_plot

growth_plot <- growth_plot + theme(axis.title.x = element_blank())
larval_duration_plot <- larval_duration_plot + theme(axis.title.x = element_blank())

# Stack the plots
Figure_1_alt <- growth_plot / larval_duration_plot / pupal_duration_plot


Figure_1_alt<- growth_plot / larval_duration_plot / pupal_duration_plot

ggsave("Figure_1_chp1.png", plot = Figure_1, width = 12, height = 5, dpi = 300)

ggsave("Figure_1_alt.png", plot = Figure_1_alt, width = 5, height = 12, dpi = 300)

```

## Fresh mass under ALAN
```{r, model wet mass, post hoc test, and figure}

#no interactions are significant so we will redo the model

model_finalwetmass <- lme(Adult.wet.mass ~ treatment + Cohort + Sex + Species,
                    random =~ 1|CohortCage,
                    data = data_mon,
                    na.action = na.omit)

summary(model_finalwetmass)
kable(summary(model_finalwetmass)$tTable)

#Run an anova
anova(model_finalwetmass, type= "marginal")
Anova(model_finalwetmass, type= "III")



# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_wetmass <- emmeans(model_finalwetmass, ~ treatment + Species + Cohort + Sex)


# Perform the pairwise comparisons with Tukey adjustment
tukey_results_wetmass <- pairs(emmeans_results_wetmass, adjust = "tukey")
print(tukey_results_wetmass)

kable(summary(tukey_results_wetmass)$Table)


```

###Figure 2A
#Plotting wet mass faceted by cohort
```{r}

fresh_mass_pred <- emmeans(model_finalwetmass, specs = ~ treatment + Species + Cohort) %>%
  as.data.frame()

fresh_mass_pred <- fresh_mass_pred %>%
  filter(!(Species == "syr" & Cohort == "1"),
         !(Species == "inc" & Cohort == "4"))


#summary of predictions
fresh_mass_pred_summary <- fresh_mass_pred %>%
  group_by(Cohort, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))





fresh_mass_plot <- ggplot(fresh_mass_pred, aes(x = treatment, y = 1000 * emmean, color = Species, group = interaction(Species, Cohort))) +
  
  # Line connecting means
  geom_line(position = position_dodge(width = 0.5)) +

  # Error bars
  geom_errorbar(aes(ymin = 1000 * lower.CL, ymax = 1000 * upper.CL), 
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Manual color for Species
  scale_color_manual(name = "Species",
                     values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Fresh mass (mg)"))) +
  xlab("Light treatment") +
  labs(tag = "A") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN"))  +

  # Clean theme styling
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),             # Remove all grid lines
    panel.border = element_blank(),           # Remove panel border
    axis.line = element_line(color = "black"),# Add axis lines
    axis.line.y.right = element_blank(),      # Remove right y-axis
    axis.line.x.top = element_blank(),        # Remove top x-axis
    strip.background = element_blank(),       # Clean facet label strip
    strip.text = element_text(size = 14),
    legend.position = c(0.25, 0.05),          # Bottom right corner
    legend.justification = c("right", "bottom"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.tag = element_text(size = 18)
     , 
    
    plot.tag.position = c(0.02, 0.98),
  )+
  
  # Facet by Cohort in a single row
  facet_wrap(~ Cohort, nrow = 1, labeller = labeller(Cohort = cohort_labels))




#plotting fresh mass again but now with the raw jittered data poitns 

fresh_mass_plot <- ggplot(fresh_mass_pred, aes(x = treatment, y = 1000 * emmean, color = Species, group = interaction(Species, Cohort))) +
    # Raw data points (behind)
  geom_jitter(
    data = data_mon,
    aes(x = treatment, y = 1000 *Adult.wet.mass, color = Species),
    alpha = 0.2,             # transparency for visibility behind means
    size = 1.8,              # small circles
    shape = 16,              # filled circles
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
  ) +
  
  # Line connecting means
  geom_line(position = position_dodge(width = 0.5)) +

  # Error bars
  geom_errorbar(aes(ymin = 1000 * lower.CL, ymax = 1000 * upper.CL), 
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Manual color for Species
  scale_color_manual(name = "Species",
                     values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Fresh mass (mg)"))) +
  labs(x = NULL) +
  labs(tag = "A") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN"))  +

  # Clean theme styling
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),             # Remove all grid lines
    panel.border = element_blank(),           # Remove panel border
    axis.line = element_line(color = "black"),# Add axis lines
    axis.line.y.right = element_blank(),      # Remove right y-axis
    axis.line.x.top = element_blank(),        # Remove top x-axis
    strip.background = element_blank(),       # Clean facet label strip
    strip.text = element_text(size = 14),
    legend.position = c(0.15, 0.70),          # Bottom right corner 
    legend.justification = c("right", "bottom"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.tag = element_text(size = 18)
     , 
    
    plot.tag.position = c(0.02, 0.98),
  )+
  
  # Facet by Cohort in a single row
  facet_wrap(~ Cohort, nrow = 1, labeller = labeller(Cohort = cohort_labels))



fresh_mass_plot

  
```

#Total wing surface area 
```{r}
data_mon <- data_mon %>%
  mutate(Total_wing_area = ifelse(Total_wing_area == 0, NA, Total_wing_area))



# Fit the model for Total Wing Area
model_totalwingarea <- lme(Total_wing_area ~ treatment + Sex + Cohort + Species + treatment*Species,
                           random = ~ 1 | CohortCage,
                           data = data_mon,
                           na.action = na.omit)

#If I remove this interaction, treatment is still significant interestingly 

# Summary of the model
summary(model_totalwingarea)
kable(summary(model_totalwingarea)$tTable)

# Run Type III ANOVA
Anova(model_totalwingarea, type = "III")
anova(model_totalwingarea, type="marginal")

# Perform Tukey post-hoc test
# Get estimated marginal means
emmeans_results_totalwingarea <- emmeans(model_totalwingarea, ~ treatment*Species + Cohort + Sex)

# Pairwise comparisons with Tukey adjustment
tukey_results_totalwingarea <- pairs(emmeans_results_totalwingarea, adjust = "tukey")
print(tukey_results_totalwingarea)


```

###Figure 2B
#Predicting and plotting total wing surface area 
```{r}
# 1. Get predicted means
wing_area_pred <- emmeans(model_totalwingarea, specs = ~ treatment + Species + Cohort) %>%
  as.data.frame()


# 1. Filter out missing combinations
wing_area_pred <- wing_area_pred %>%
  filter(!(Species == "syr" & Cohort == "1"),
         !(Species == "inc" & Cohort == "4"))

# 3. Create plot
wing_area_plot <- ggplot(wing_area_pred, aes(x = treatment, y = emmean, color = Species, group = interaction(Species, Cohort))) +
  
  # Line connecting means
  geom_line(position = position_dodge(width = 0.5)) +

  # Error bars
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Manual color for Species
  scale_color_manual(name = "Species",
                values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Total wing area (mm"^2, ")"))) +
  xlab("Light treatment") +
  labs(tag = "B") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
  # Facet by Cohort in a single row with custom labels
  facet_wrap(~ Cohort, nrow = 1, labeller = labeller(Cohort = cohort_labels)) +

  # Clean theme styling
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    axis.line.y.right = element_blank(),
    axis.line.x.top = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    legend.position = c(0.95, 0.05),
    legend.justification = c("right", "bottom"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.tag = element_text(size = 18)
  )


#plotting again now but with the raw data points 
# 3. Create plot
wing_area_plot <- ggplot(wing_area_pred, aes(x = treatment, y = emmean, color = Species, group = interaction(Species, Cohort))) +
geom_jitter(
  data = dplyr::filter(data_mon, Total_wing_area != 0),
  aes(x = treatment, y = Total_wing_area, color = Species),
  alpha = 0.2,
  size = 1.8,
  shape = 16,
  position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
) +
  # Line connecting means
  geom_line(position = position_dodge(width = 0.5)) +

  # Error bars
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.1, position = position_dodge(width = 0.5)) +
  
  # Mean points
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  
  # Manual color for Species
  scale_color_manual(name = "Species",
                values = c("inc" = "deepskyblue4", "syr" = "red3"),
                     labels = c(expression(italic("A. incarnata")),
                                expression(italic("A. syriaca")))) +
  
  # Axis labels
  ylab(expression(paste("Total wing area (mm"^2, ")"))) +
  xlab("Light treatment") +
  labs(tag = "B") +
  scale_x_discrete(labels = c("ML" = "Control", "AL" = "ALAN")) +
  
  # Facet by Cohort in a single row with custom labels
  facet_wrap(~ Cohort, nrow = 1, labeller = labeller(Cohort = cohort_labels)) +

  # Clean theme styling
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    axis.line.y.right = element_blank(),
    axis.line.x.top = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    legend.position = c(0.15, 0.03),
    legend.justification = c("right", "bottom"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.tag = element_text(size = 18)
  )

wing_area_plot

```

###Figure 2
#Putting together fresh wet mass and total wing surface area plots together 
```{r}

Figure_2<- fresh_mass_plot / wing_area_plot

# Save the combined plot as a PNG
ggsave("Figure_2_chp1.png", plot = Figure_2, width = 13, height = 9, dpi = 300)

```

#Dry mass model 
```{r}
# No interactions are significant, so we will redo the model using dry mass
model_finaldrymass <- lme(Adult.dry.mass ~ treatment + Cohort + Sex + Species,
                          random = ~ 1 | CohortCage,
                          data = data_mon,
                          na.action = na.omit)

# Summary of the model
summary(model_finaldrymass)
kable(summary(model_finaldrymass)$tTable)

# Run a Type III ANOVA
anova(model_finaldrymass, type= "marginal")
#Anova(model_finaldrymass, type = "III")
#don't know why I am getting an error for this 

# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_drymass <- emmeans(model_finaldrymass, ~ treatment + Species + Cohort + Sex)

# Pairwise comparisons with Tukey adjustment
tukey_results_drymass <- pairs(emmeans_results_drymass, adjust = "tukey")
print(tukey_results_drymass)

# Output results as a table
kable(summary(tukey_results_drymass)$summary)

```

###Figure 3A
#correlation between fresh and dry mass 
```{r}

# Fit the ANCOVA model
model_drymass <- lm(Adult.dry.mass ~ Adult.wet.mass * treatment + Sex + Species + Cohort, data = data_mon)

# Summary of the model
summary(model_drymass)
print(summary(model_drymass), digits = 4)


# Optionally format the summary with kable
library(knitr)
kable(summary(model_drymass)$coefficients, digits = 3)

#calculating the means so they can be plotted 
means <- data_mon %>%
  group_by(treatment) %>%
  summarise(
    mean_wet = mean(Adult.wet.mass, na.rm = TRUE),
    mean_dry = mean(Adult.dry.mass, na.rm = TRUE)
  )

#Time to plot
fresh_dry_plot <- ggplot(data_mon,
                         aes(x = Adult.wet.mass*1000,
                             y = Adult.dry.mass*1000,
                             color = treatment)) +
  geom_jitter(width = 0.1, height = 0.1, size = 1.8, alpha = 0.4) +
  scale_color_manual(name = "Light treatment",
                     values = c("ML" = "#004B74", "AL" = "#E69F00")
,
                     labels = c("Control", "ALAN")) +
geom_smooth(aes(color = treatment), method = "lm", se = FALSE, size = 1,
            linetype = "solid") +
  labs(
    x = "Fresh mass (mg)",
    y = "Dry mass (mg)",
    color = "Treatment"
  ) +
    labs(tag = "A") +
  
  theme_minimal() +
  theme(
    legend.position = c(0.3, 0.9),,
    text = element_text(size = 14),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", size = 0.8)
  )


fresh_dry_plot

#modifying the axis point 
fresh_dry_plot <- ggplot(data_mon,
                         aes(x = Adult.wet.mass*1000,
                             y = Adult.dry.mass*1000,
                             color = treatment)) +
  geom_jitter(width = 0.1, height = 0.1, size = 1.8, alpha = 0.4) +
  scale_color_manual(
    name = "Light treatment",
    values = c("ML" = "#004B74", "AL" = "#E69F00"),
    labels = c("Control", "ALAN")
  ) +
  geom_smooth(aes(color = treatment), method = "lm", se = FALSE,
              size = 1, linetype = "solid") +
  labs(
    x = "Fresh mass (mg)",
    y = "Dry mass (mg)"
  ) +
  labs(tag = "A") +
scale_x_continuous(
  limits = c(310, NA),                 # 👈 x-axis starts at 350
  expand = expansion(mult = c(0, 0.02))  # optional small padding
) +

  theme_minimal() +
  theme(
    legend.position = c(0.28, 0.91),
    text = element_text(size = 14),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", size = 0.8)
  )

```

#Forewing length model
```{r}

model_length <- lme(Forewing_length_mm ~ treatment + Cohort + Sex + Species,
                          random = ~ 1 | CohortCage,
                          data = data_mon,
                          na.action = na.omit)

summary(model_length)
kable(summary(model_length)$tTable)


# Run a Type III ANOVA
anova(model_length, type="marginal")
Anova(model_length, type = "III")

# Perform Tukey post-hoc test
# Get the estimated marginal means
emmeans_results_length <- emmeans(model_length, ~ treatment + Species + Cohort + Sex)

# Pairwise comparisons with Tukey adjustment
tukey_results_length <- pairs(emmeans_results_length, adjust = "tukey")
print(tukey_results_length)

# Output results as a table
kable(summary(tukey_results_length)$summary)


```

###Figure 3B
#Predicting and plotting for forewing length 
```{r}
forewing_pred <- emmeans(model_length, specs = c("treatment", "Species")) %>%
  as.data.frame()


# Summary of predictions
forewing_pred_summary <- forewing_pred %>%
  group_by(Species, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))



length_plot<- ggplot(forewing_pred_summary, aes(x = Species, y = emmean, group = treatment)) + 
      # Raw data points (behind)
  geom_jitter(
    data = data_mon,
    aes(x = Species, y = Forewing_length_mm, color = Treatment),
    alpha = 0.2,             # transparency for visibility behind means
    size = 1.8,              # small circles
    shape = 16,              # filled circles
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
  ) +
  
  geom_line(position = position_dodge(width = 0.5)) +
  geom_point(aes(color = treatment), size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymax = upper.CL, ymin = lower.CL, color = treatment), 
                width = 0.1, position = position_dodge(width = 0.5)) +
  scale_color_manual(name = "Light treatment",
                     values = c("ML" = "#004B74", "AL" = "#E69F00")
,
                     labels = c("Control", "ALAN")) +
  ylab(expression(paste("Forewing length (mm)"))) +
  xlab("Species") +
  labs(tag = "B") +
  scale_x_discrete(labels = c("inc" = expression(italic("A. incarnata")),
                              "syr" = expression(italic("A. syriaca")))) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),             # Remove all grid lines
    panel.border = element_blank(),           # Remove border
    axis.line = element_line(color = "black"),
    axis.line.x.top = element_blank(),        # Remove top x-axis
    axis.line.y.right = element_blank(),      # Remove right y-axis
    legend.position = "none",          # Top-left corner
    legend.justification = c("left", "top"),
    legend.background = element_blank(),      # No outline
    legend.key = element_blank(),
    plot.tag = element_text(size = 16)
  )

length_plot
#made sure to use color blind friendly color scale 


```


#Wing loading 
```{r}
data_mon <- data_mon %>%
  mutate(Wingloading = ifelse(Wingloading == 0, NA, Wingloading))


# Fit the model for Wingloading
model_wingloading <- lme(Wingloading ~ treatment + Sex + Cohort + Species,
                         random = ~ 1 | CohortCage,
                         data = data_mon,
                         na.action = na.omit)

# Summary of the model
summary(model_wingloading)
kable(summary(model_wingloading)$tTable)

# Run Type III ANOVA
anova(model_wingloading, type = "marginal")
Anova(model_wingloading, type = "III")

# Perform Tukey post-hoc test
# Get estimated marginal means
emmeans_results_wingloading <- emmeans(model_wingloading, ~ treatment*Species + Cohort + Sex)

# Pairwise comparisons with Tukey adjustment
tukey_results_wingloading <- pairs(emmeans_results_wingloading, adjust = "tukey")
print(tukey_results_wingloading)


```

###Figure 3C
#Plotting wing loading 
```{r}
wingload_pred <- emmeans(model_wingloading, specs = ~ treatment + Cohort) %>%
  as.data.frame()

# Summary of predictions
wingload_pred_summary <- wingload_pred %>%
  group_by(Cohort, treatment) %>%
  summarize(emmean = mean(emmean), 
            lower.CL = mean(lower.CL), 
            upper.CL = mean(upper.CL))


wingloading_plot<- ggplot(wingload_pred_summary, aes(x = Cohort, y = emmean, group = treatment)) + 
    geom_jitter(
      data = data_mon,
      aes(x = Cohort, y = Wingloading, color = Treatment),
      alpha = 0.2,
      size = 1.8,
      shape = 16,
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5)
    ) +
    geom_line(position = position_dodge(width = 0.5)) +
    geom_point(aes(color = treatment), size = 3, position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymax = upper.CL, ymin = lower.CL, color = treatment),
                  width = 0.1, position = position_dodge(width = 0.5)) +
    scale_color_manual(name = "Light treatment",
                       values = c("ML" = "#004B74", "AL" = "#E69F00"),
                       labels = c("Control", "ALAN")) +
    ylab(expression(paste("Wing loading (mg/mm"^2, ")"))) +
    xlab("Cohort") +
    labs(tag = "C") +
    
    # ✅ REMOVE 0 LABEL
  scale_y_continuous(
  labels = function(x) ifelse(x == 0, "", x),
  expand = expansion(mult = c(0.05, 0.15))  # 5% bottom, 15% top padding
) +

    theme_bw(base_size = 16) +
    theme(
      panel.grid = element_blank(),
      panel.border = element_blank(),
      axis.line = element_line(color = "black"),
      axis.line.x.top = element_blank(),
      axis.line.y.right = element_blank(),
      legend.position = "none",
      legend.justification = c("left", "top"),
      legend.background = element_blank(),
      legend.key = element_blank(),
      plot.tag = element_text(size = 16)
    )

wingloading_plot
#plotting it to see what it would be like if we plotted based on species 



```

###Figure 3
#putting together Figure 3 
```{r}
Figure_3<- fresh_dry_plot + length_plot + wingloading_plot

ggsave("Figure_3_chp1.png", plot = Figure_3, width = 13, height = 5, dpi = 300)


```

#Path analysis 
```{r}

#cleaning the did to get rid of all of the na's 
data_mon_clean <- data_mon %>%
  filter(!is.na(Sex))


data_mon_clean$mstatus <- ifelse(as.character(data_mon_clean$Cohort) %in% c("1", "2"), "Group_1_2",
                          ifelse(as.character(data_mon_clean$Cohort) %in% c("3", "4"), "Group_3_4", NA))

data_mon_clean$mstatus <- factor(data_mon_clean$mstatus, levels = c("Group_1_2", "Group_3_4"))


data_mon_path <- data_mon_clean %>%
  dummy_cols(select_columns = c("mstatus", "Species", "Sex"),
             remove_first_dummy = TRUE,
             remove_selected_columns = TRUE)

#standardize the varibales so there isn't such a wide variety in effect sizes/ estimates 
data_mon_path$growth.rate <- scale(data_mon_path$growth.rate)
data_mon_path$Larval_duration <- scale(data_mon_path$Larval_duration)
data_mon_path$Pupal_duration <- scale(data_mon_path$Pupal_duration)
data_mon_path$Adult.wet.mass <- scale(data_mon_path$Adult.wet.mass)
data_mon_path$Total_wing_area <- scale(data_mon_path$Total_wing_area)



#path
path_diagram <- '
growth.rate ~ treatment + mstatus_Group_3_4 + Species_syr
Larval_duration ~ growth.rate + treatment + mstatus_Group_3_4 + Species_syr
Pupal_duration ~ growth.rate + Larval_duration + treatment + mstatus_Group_3_4 + Species_syr + Sex_M
Adult.wet.mass ~ growth.rate + Larval_duration + Pupal_duration + treatment + mstatus_Group_3_4 + Species_syr + Sex_M
Total_wing_area ~ growth.rate + Larval_duration + Pupal_duration + Adult.wet.mass +  treatment + mstatus_Group_3_4 + Species_syr + Sex_M
'



fit <- sem(path_diagram, data = data_mon_path)


summary(fit, fit.measures = TRUE, standardized = TRUE, rsquare=T)




summary(fit, fit.measures = TRUE, standardized = TRUE)


```

#Kaplan Survival Curve
```{r}
# Calculate survival time in days
data_mon$survival_time <- as.numeric(difftime(data_mon$Date.of.death, data_mon$Hatch.Date, units = "days"))

# Create an event indicator (assuming death is coded as 1 and censoring as 0)
data_mon$event <- ifelse(!is.na(data_mon$Date.of.death), 1, 0)


# Create survival object
surv_object <- Surv(time = data_mon$survival_time, event = data_mon$event)

# Fit Kaplan-Meier survival curve by treatment
km_fit <-survfit(surv_object ~ treatment, data = data_mon)


# Plot Kaplan-Meier curves
survival_plot<- ggsurvplot(
  km_fit, 
  data = data_mon,
  pval = TRUE,                # Show p-value for log-rank test
  conf.int = TRUE,            # Show confidence interval
  risk.table = TRUE,          # Display risk table
  xlab = "Days",
  ylab = "Survival Probability",
  legend.title = "Treatment",
  legend.labs = c("Control", "ALAN"),  # Customize legend labels
  palette = c("royalblue4", "goldenrod1") # Set colors for AL and ML
)


ggsave("survival_plot_chp1.png", plot = survival_plot$plot, width = 10, height = 6, dpi = 300)

# Save the full survival plot including risk table
ggsave("survival_plot_with_risktable.png",
       plot = print(survival_plot),  # This forces full rendering including table
       width = 10,
       height = 8,
       dpi = 300)



# Log-rank test
survdiff(surv_object ~ treatment, data = data_mon)
```

### Figure S5
#correlation plot 
```{r}
data_ML %>%
  dplyr::select (c(growth.rate, Larval_duration, Pupal_duration, Adult.wet.mass, Total_wing_area)) %>%
  ggpairs()


data_mon <- data_mon %>%
  mutate(Treatment = factor(Treatment, levels = c("ML", "AL")))


sup_fig<- data_mon %>%
  dplyr::select(
    Treatment,
    growth.rate,
    Larval_duration,
    Pupal_duration,
    Adult.wet.mass,
    Total_wing_area
  ) %>%
  # Rename columns
  dplyr::rename(
    `Light treatment` = Treatment,
    `Growth rate` = growth.rate,
    `Larval duration` = Larval_duration,
    `Pupal duration` = Pupal_duration,
    `Fresh mass` = Adult.wet.mass,
    `Total wing area` = Total_wing_area
  ) %>%
  # Rename treatment levels
  mutate(`Light treatment` = factor(`Light treatment`, 
                                    levels = c("ML", "AL"),
                                    labels = c("Control", "ALAN"))) %>%
  ggpairs(
    aes(color = `Light treatment`, fill = `Light treatment`),
    upper = list(continuous = "cor"),             # correlation numbers
    lower = list(continuous = wrap("points", alpha = 0.5)),  # transparent points
    diag  = list(continuous = wrap("densityDiag", alpha = 0.5)) # transparent densities
  ) +
  scale_color_manual(values = c("Control" = "#004B74", "ALAN" = "#FFD500")) +
  scale_fill_manual(values = c("Control" = "#004B74", "ALAN" = "#FFD500")) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  # Make correlation numbers match treatment colors
  ggplot2::theme(
    strip.text = element_text(color = "black")
  )

ggsave("sup_fig.png", plot = sup_fig, width = 12, height = 7, dpi = 300)
```


###Figure S6
#looking at caterpillar location 
```{r}

# load location data 
location_data<- read_sheet("https://docs.google.com/spreadsheets/d/19eTdjvJmQ6QqMTdKgQp42JZcs2Znglq00gkSpDcbC5I/edit?gid=0#gid=0")


location_data <- location_data[, !(names(location_data) %in% "Time")]



data_mon_small <- data_mon %>%
  filter(Cohort == 1) %>%      # keep only Cohort 1
  dplyr::select(
    Individual, Treatment, Cohort, Cage, Species,
    Day.7.date, Day.7.node, Day.7.avaiable.nodes,
    Ratio.location, Day.7.Leaf.orientation
  )


combined_data <- data_mon %>%
  dplyr::select(
    Individual, Treatment, Cohort, Cage, Species,
    Day.7.date, Day.7.node, Day.7.avaiable.nodes,
    Ratio.location, Day.7.Leaf.orientation
  ) %>%
  left_join(location_data, by = "Individual") %>%
  mutate(
    Observation_type = if_else(is.na(Observation_type), "Afternoon", Observation_type)
  )

combined_data <- dplyr::select(combined_data,
  -`Day.7.node.y`,
  -`Cohort.y`,
  -`Species.y`,
  -`Cage.y`,
  -`Treatment.y`,
  -`Day.7.Leaf.orientation.y`
) %>%
  # rename the .x columns to clean names
  rename(
    Day.7.node = `Day.7.node.x`,
    Cohort = `Cohort.x`,
    Cage = `Cage.x`,
    Species = `Species.x`,
    Treatment = `Treatment.x`,
    Day.7.Leaf.orientation = `Day.7.Leaf.orientation.x`
  ) %>%
    mutate(Day.7.node = as.numeric(Day.7.node))


# Calculate summary stats (mean and SE)
summary_data <- combined_data %>%
  group_by(Observation_type, Treatment) %>%
  summarise(
    mean_node = mean(Day.7.node, na.rm = TRUE),
    se_node   = sd(Day.7.node, na.rm = TRUE)/sqrt(n()),
    .groups = "drop"
  )

library(ggplot2)




cat_location_plot<- ggplot(combined_data, aes(x = factor(Observation_type, levels = c("Dawn", "Afternoon", "Dusk")), 
                          y = Day.7.node, color = Treatment)) +
  # Raw data points, jittered to match mean/error bars
  geom_jitter(aes(color = Treatment), 
              position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.3), 
              alpha = 0.5, size = 2) +
  # Mean points
  geom_point(data = summary_data, 
             aes(x = factor(Observation_type, levels = c("Dawn", "Afternoon", "Dusk")), 
                 y = mean_node, color = Treatment), 
             position = position_dodge(width = 0.3), size = 3) +
  # Error bars for SE
  geom_errorbar(data = summary_data, 
                aes(x = factor(Observation_type, levels = c("Dawn", "Afternoon", "Dusk")), 
                    y = mean_node, ymin = mean_node - se_node, ymax = mean_node + se_node, color = Treatment),
                position = position_dodge(width = 0.3), width = 0.2) +
  # Flip Y-axis
  scale_y_reverse() +
  # Custom colors
  scale_color_manual(name = "Light treatment",
                     values = c("ML" = "#004B74", "AL" = "#E69F00"),
                     labels = c("Control", "ALAN")) +
  labs(
    x = "Time of day",
    y = "Caterpillar location (nodes from the top)",
    color = "Treatment"
  ) +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    legend.position = c(0.04, 0.04),   # bottom-left corner
    legend.justification = c(0, 0),    # anchor legend by bottom-left
    legend.background = element_rect(fill = "transparent") # optional: make background transparent
  )


ggsave("cat_location_plot.png", plot = cat_location_plot, width = 8, height = 5, dpi = 300)

```


